### 2013-08-22 SingularStochasticControl.tex

\documentclass{article}

\usepackage{amsmath}
\usepackage{amscd}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amsthm}

\usepackage{circuitikz}
\usepackage{pgf}
\usepackage{tikz}
\usetikzlibrary{arrows,snakes,backgrounds}
% \usetikz
\usepackage{subfig}

\usepackage[super]{nth}
% \usepackage{appendix}
% \usepackage{listings}
% \usepackage{color}

\usepackage{hyperref}
%\usepackage{url}

\usepackage{cleveref}

\usepackage{aviolov_style}
\usepackage{local_style}

\begin{document}

\title{Singular Stochastic Control} \author{Alexandre Iolov 
$<$\href{mailto:aiolo040@uottawa.ca}
		{aiolo040 at uottawa dot ca}$>$}

\date{\today}

\maketitle

\abstract{How to solve HJB equations, whose Hamiltonian is linear in the
control and there is no guarantee that $\di_x v \neq 0$?}

\section{Problem Statement}

Consider a 2nd order HJB equation in 1-d with no running costs and fixed
terminal time $\tf$:
\begin{equation}
\begin{gathered}
dX_s =  (\a(s) - X_s) \intd{s} + \b \intd{W_s} 
\\
J[\a] = M(X_\tf)
\\
\di_t v(x,t) + \inf_{\a} \left\{ \tfrac{\b^2}{2} \di_x^2 v + 
(\a(t) - x) \di_x v \right\} = 0
\\
\ldots + \textrm{BCs, TCs}
\end{gathered}
\label{eq:HJB_singular}
\end{equation} 
We are solving $v(x,t)$ on some closed, bounded domain $\O \subset \R \times
[0,\tf]$.

As long as $\di_x v \neq 0$, we will have a 'bang-bang' type control for
$\a(t)$:
\begin{equation}
\a(t) =
\begin{cases}
\amin & \di_x v > 0
\\
\amax & \di_x v < 0
\end{cases}
\end{equation}
But what to do if $\di_x v= 0$?

There is a solution for this problem in the deterministic optimal control
theory. In the deterministic optimal control, the quantity:
$$
H(x,p) = \inf_{\a} \left\{
(\a(t) - x) p \right\}
$$
is called the 'Hamiltonian'. The problem is called 'singular', when $p=0$, since
then it is impossible to know the value of $\a$ that attains the minimum. The
remedy is to note that along optimal trajectories, $x^*,p^*$ Pontryagin's
Maximum Principle is satisfied, $x,p$ form a Hamiltonian Dynamical System and
the Hamiltonian is an integral of motion - $H = \const$. Thus along optimal
trajectories the full time derivative of $H$ is zero, $d_t H=0$. Thus we can
differentiate $H$ until $\a$ props up again and then set the resulting
expression to zero to obtain an equation for $\a$.

But is there an analogue to this in the Stochastic Case?

For our very simple problem with no running costs, the stochastic Hamiltonian
is simply:
$$
H(x,p,q) = \inf_{\a} \left\{\tfrac{\b^2}{2} q +
(\a(t) - x) p \right\}
$$

\bibliographystyle{plain}
\bibliography{library,local}
% \bibliography{local}

\end{document}