### 2013-08-22 draft.tex

% \documentclass[12pt]{iopart}
\documentclass{article}

\usepackage{amsmath}
\usepackage{amscd}
\usepackage{amssymb}
\usepackage{amsfonts} 
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amsthm}

\usepackage{tikz}
\usetikzlibrary{arrows,snakes,backgrounds}
% \usepackage{circuitikz}
% \usepackage{pgf}
% \usetikz
\usepackage{subfig}

\usepackage{algpseudocode}
\usepackage{algorithm}

\usepackage[super]{nth}
\usepackage{appendix}
% \usepackage{listings}
\usepackage{color}

\usepackage{hyperref}
%\usepackage{url}

\usepackage{cleveref}
\usepackage{cancel}
\usepackage{slashbox}

\usepackage{aviolov_style}
\usepackage{local_style}

\newtheorem{thm}{Theorem}[section]
\newtheorem{lemma}{Theorem}[thm]
% \theoremstyle{definition}
\newtheorem{ex}{Example}[thm]
\newtheorem{defn}{Definition}[thm]

% \def\includegraphic{}
% \def\includegraphics{}

\begin{document}

\title{Stochastic Optimal Control of Single Neuron Spike Trains
\\
---
\vskip5pt
Draft
}
\author{Alexandre Iolov, Susanne Ditlevsen, Andr\'e Longtin,
		}		
% \address{Department of Mathematics and Statistics, University of Ottawa, Ottawa, Canada 
% \\ Department of Mathematical Sciences, University of Copenhagen, Copenhagen,
% Denmark }
% \ead{$<$\href{mailto:aiolo040@uottawa.ca}
% 		{aiolo040 at uottawa dot ca}$>$}

\date{\today}

\maketitle

\abstract{
{\em Objective.} External control of exact spike timings in single neurons is
important for basic understanding of neuronal activity, and has
potential for improving neural prostheses and in medical applications
related to neuronal disorders, such as Parkinson's disease or epilepsy.
The goal of this paper is the design of optimal controlled
electrical stimulation patterns to a cell to achieve a target
spike train under physiological constraints not to damage tissue and
minimize possible side effects.
{\em Approach.} We pose an optimal control problem to precisely specify
the spike timing in a noisy leaky integrate-and-fire model of a
neuron. In particular, we allow for the noise to be of arbitrary intensity. The optimal control problem is solved using Dynamic Programming if the controller has access to the voltage (closed-loop control) and using a Maximum Principle for the transition density if the controller only has access to the spike timings (open loop
control).
{\em Main results.} We have developed a stochastic optimal control
scheme to obtain precise spike timings, which is not limited
by the intensity of the noise in the system. Given this scope, the
numerical demands are non-trivial, but simulations show that the
algorithms are potentially feasible. It however remains to see if
they can be successfully applied in a real-time experimental setting.
Furthermore, we show how wrong the control strategy can perform if we
assume that the noise is negligible.
{\em Significance.} More precise targeting and control of neural
activity to achieve optimal performance in a noisy environment, in
which most neurons are embedded, have great potential in health
care. The main contribution is the online feedback control of a noisy
system through
modulation of the input, taking into account physiological constraints
on the control, as well as incurred costs to diminish adverse effects.
}

\tableofcontents

MSC KEYS: 49N90 90C39 93E20 93C20 93C95

\section{Introduction}
Manipulation of individual neurons through external electrical
stimulation provides a means to
controlling the spiking activity. Applications may arise, for
example, in brain-machine interfaces or in artificial prosthetics.
A natural goal is to make a cell produce a pre-specified spike train,
under certain constraints on the modulation of the input.
We consider this problem in the framework of Stochastic Optimal Control and
give both a feedback solution (closed-loop), when the cell voltage is
explicitly observable, as well as an open-loop solution when only the
occurrence of spikes is observable. Importantly, we allow for an
arbitrary noise intensity.

Both theoretically and in practice related problems have been
addressed in the literature. One objective has been to
obtain either minimum or maximum interspike intervals lengths,
when the input is constrained to be between some prespecified upper
and lower bounds, see e.g. \cite{Lee1994,Lefebvre1987} for a
mathematical treatment, or \cite{Danzl2009,Nabi2012,Wu2009} in a neuronal
context. Another objective is to break a pathological synchronous
firing pattern in clusters of neurons, highly relevant for neurological
disorders such as epilepsy and Parkinson's disease,
\cite{Nabi2013a,Nabi2011}, see also \cite{Feng2007b}.

Our objective, namely targeting exact spike times in single neurons, has been
considered mainly in the open-loop context, and in either absence of or for
small noise. In \cite{Ahmadian2011} they use the Spike Response Model,
\cite{Gerstner2002}, to control output target spike trains and implement their
scheme on pyramidal cells in mouse cortical slices. Their method is numerically
efficient and allows for the simultaneous control of many neurons. However the
efficiency of the numerics strongly relies on the assumption that the diffusion
noise, the value of $\b$ in \cref{eq:X_evolution_uo} below, is small enough, see
the discussion above and below equation 12 in \cite{Ahmadian2011}. They only
work with open-loop control. The difference between the objective in
\cite{Ahmadian2011} and what we consider is that they maximize the probability
of spiking at some given time $t^*$, whereas we minimize the mean squared
difference between the realized spike time, $T_{sp}$ and the desired, $t^*$.
Moehlis et al. \cite{Moehlis2006} work with a Phase Response Model to obtain
spikes at exact times, while keeping the root mean square of the input to a
minimum. A similar approach is taken in \cite{Dasanayake2011}. They do not
consider noise, though. In \cite{Nabi2013}, the methods are implemented on brain
slices of pyramidal neurons of rat hippocampus. [[[ALEX, Is it closed- or open
loop? Could you SHORTLY describe the difference of what they do and what we
do?]]] \{ALEX: They use Phase Response Curves, assuming that a spike coincides
with $\th=0$; no noise, Pontryagin Maximum Principle, solved numerically $=>$
open-loop. Short Description: Since they do not consider noise, it is
straight-forward to make the neuron spike when you want it, and in fact that can
be done in many ways. Thus their optimization reduces to finding the way to make
the neuron spike at a pre-specified time, using least energy (the integral of
$u^2$.\}. While the Phase Response Curve Model is a parsimonious and effective
way to describe a neuron's response to a stimulus, it is only valid in the
supra-threshold regime, when the unstimulated neuron is periodically spiking.
In \cite{Feng2003} they use a leaky integrate-and-fire model and also
use open-loop control, but to obtain exact spike timings, they choose an
objective of minimizing the variance of the membrane potential at the desired
spike time, while forcing the mean of the membrane potential at this time point
to equal the threshold. This provides exact solutions since it does not involve
first-passage times, but has the drawback of getting a non-negligible
probability of being way-off the objective.

Our objective of imposing a certain timing sequence of the spike train using an
externally applied control is obtained in both the closed- and
open-loop setting, and we specifically include the noise in the
calculations of the controls. We restrict the controlled input to stay
within pre-specified bounds, and also include a cost function to
minimize intervention.
Our main contributions are... (((what we have done which has not been
done before, numerical merits...)))
\{ALEX: Basically, we deal with a significant noise component. Everything
else follows, i.e. the objective only makes sense if the ISI is a random variable.
The PDEs (HJB and Max Principle) only occur because of the noise
component, o/w we would have ODEs. The numerical schemes are simple given the
PDEs, we just use the Crank-Nicholson method for the PDEs and a basic
gradient-descent for the optimization in the open-loop case. The numerics take a
non-trivial time to run - hard to see them coming down to real-time, i.e.
10-20 ms, \ldots but possible. We also explore a wide spectrum of parameter
values, in particular we do  not restrict our attention to the
autonomously-spiking, supra-threshold regime.\}

[[[Susanne, we should mention Whittle \cite{Whittle1996}, but here, or together
with other monographs on optimal control?]]] \{ALEX: Indeed we should mention
Whittle, in particular mention that he discusses analytically the problem in the
closed-loop case (only), but not the open-loop case. He does not provide
numerics in either case. I would refrain from citing other books on
stochastic optimal control, here. Books commonly cited are
\cite{Fleming1975,Krylov2008} and the chapter on HJB equation in Oksendal,
\cite{Oksendal2007}, some of them are cited in later sections\ldots\}

The paper is structured as follows: First we describe the neuron model and
formalize the control objective. Then we describe a feedback-based solution,
which assumes that the controller has detailed access to the voltage trajectory.
Then we relax the observation assumption so that the controller only has access
to the spike timings. Finally, we compare the two methods through
simulations against a simple-minded control technique which ignores
the stochastic input to the neuron.

\section{Problem Formulation}
A basic but useful model for the neural membrane potential evolution is the
noisy leaky-integrate-and-fire model:
\begin{equation}
\begin{gathered} 
dX_s = \left(\Iext(t) - \frac{X_s}{\tc} \right) \intd{s} + \b \intd{W_s},
\\
X(0) = 0,
\\
X(\ts) = \xth \implies   
\begin{cases}
X(\ts^+) = 0 &  
\end{cases}
\end{gathered}
\label{eq:X_evolution_uo}
\end{equation}
Here $X$ represents the membrane electric potential which tries to revert to $0$
with a time-const of $\tc$, $dW$ is a Brownian motion increment scaled by $\b$
and $\Iext(t)$ is the deterministic external input to the cell. Having last
spiked at time $0$, the potential hits $\xth$ at some time $\ts$, the potential
resets to $0$ and we start all over again. We write $\ts^+$ in the formula for
the reset since we are working in continuous time, i.e.\ it is really the
voltage's limit from the right at $\ts$ that is 0. For our purposes, we will
always have that the threshold is set to one, $\xth = 1$.

Suppose that we have some control over the external current such that it can be
decomposed as
\begin{equation}
\Iext(t) = \m + \a(t)
\label{eq:current_mu_alpha}
\end{equation}
where $\m$ is an uncontrollable, but constant part of the external current and
$\a(t)$ is controllable, i.e.\ it can be chosen to achieve some goal. 
A natural goal is to attempt to control the spike time, $\ts$. 
That is we would like to ask how do we choose $\a(t)$ such that $\ts = \T$.
A natural optimal control objective to achieve this is the least squares solution
\begin{equation}
\a(t) = \argmin_{\a(t)} \{\, \Exp[(\ts- \T)^2] \,\}
\label{eq:OC_LS_variance}   
\end{equation}
where expectation is taken with respect to the distribution of the trajectories
of $X_t$.

Often the control has certain constraints. The most common are simple box
constraints:
\begin{equation}
\a(t) \in [\amin, \amax] \quad \forall t
\label{eq:bound_constraints_alpha}
\end{equation}

% \subsection{Optimal Control Formulation}
In addition to \cref{eq:OC_LS_variance}, we will also add to our objective a
running cost based on the control. This regularizes the problem eliminating the
subtleties of singular-control situations and it might be argued appropriate to
avoid excessive control. WE SHOULD DISCCUSS THIS IN GREATER DETAIL, BUT PERHAPS
LATER. (ALEX:\cite{Ahmadian2011} state that they impose a similar control
penalty to avoid excessive charge building up on the cell.)

So we seek an optimal control, $\a^*$, that solves
\begin{align}
J[\a(\cdot)] =&  
\Exp\left[
\e \int_0^\ts  \a^2(s) \intd{s}
+  
(\ts - \T \big)^2 \right]  
\label{eq:OC_LS_variance_energy}   
\\ \notag
\a^*(\cdot) =& \argmin_{\a(\cdot)} J[\a(\cdot)]. 
\end{align}
where $\e$ measure how much weight we put on minimizing the energy expenditure.
If $\e = 0$, then we do not care at all about the expended
energy.

It will often be the case that we write $\a$ and mean either a function or a
value of that function at a particular time. And that this function could be
random, i.e.\ a stochastic process. Since $\a$ is used to control the behaviour
of $X$, it will be at least measurable with respect to the filtration of $X$,
see \cite{Krylov2008}. We will try to make clear below when we are refering to
$\a$ as a stochastic process and when we are merely refering to its particular
value at some time. For example, in \cref{eq:OC_LS_variance_energy}, $\a(\cdot)$
refers to the function, while $\a(s)$ refers to that function's value, possibly
random, at time $s$.

We will consider two control contexts -- {\sl closed-loop} and  {\sl open-loop}
control. In closed-loop control, the value of $X_t$ at time $t$ is observable
and can be used in determining the control. In open-loop control only the spike
times are observable. In the closed-loop context, we might better write the
control as $$\a = \a(x,t)$$ to indicate its dependence on $X_t$ and to express
that it will be updated based on the time-course of $X$. 
In the open-loop context, we would instead write $$\a = \a(t)$$ to indicate that
$\a(\cdot)$ is decided for all times at time 0. 
The techniques used in the two scenarios will be different. For the closed-loop
scenario we will use Dynamic Programing, \cite{Fleming1975}, in order to obtain
the optimal control, while for the open-loop scenario we will use a form of the
Maximum Principle applied on the transition density of the controlled process,
\cite{Ahmed1981,Borzi2012}.

Crucially, we assume that the model parameters, $\m,\tc,\b$ in
\cref{eq:X_evolution_uo} are known.

\subsection{Parameter Regimes}
Different  parameter regimes can be envisioned given \cref{eq:X_evolution_uo},
depending on whether the noise intensity, $\b$, is
relatively high or low, and whether the external, uncontrollable bias current,
$\m$, induces spikes in the absence of noise or not.

Spikes will occur in the absence of noise, if and only
if $$ \m > \frac 1\tc,$$ which is called the supra-threshold regime. When $\m
\leq 1/\tc$, the regime is called the sub-threshold regime.
In addition we will impose two values of $\b$, which we
will call high-noise and low-noise, respectively.
Example values for each parameter regime are given in \cref{tab:regimes}, and we
visualize a single path for each regime in \cref{fig:regime_path_examples}.
\begin{table}
\begin{tabular}{|l||{c}|{c}|}
\hline
\backslashbox{$\m$}{$\b$}
& $1.5$ & $0.3$ \\
\hline
$1.5 / \tc $ &Supra-threshold-High-noise & Supra-threshold-Low-noise \\ 
\hline
$0.1 / \tc$   &Sub-threshold-High-noise & Sub-threshold-Low-noise \\
\hline
\end{tabular}
\caption{Regime labels and example values. Note that for the numerical
experiments below, we use $\tc = 0.5$}
\label{tab:regimes}
\end{table}
%\usepackage{graphics} is needed for \includegraphics
\begin{figure}[htp]
\begin{center} 
  \includegraphics[width=0.99\textwidth]{Figs/PathSimulator/path_T=14_combined.pdf}
  \caption[labelInTOC]{Example Trajectories from \cref{eq:X_evolution_uo}
  using the parameter values from \cref{tab:regimes}. A)
  Supra-Threshold-High-Noise B) Supra-Threshold-Low-Noise C) Sub-Threshold-High-Noise D) Sub-Threshold-Low-Noise}
  \label{fig:regime_path_examples} 
\end{center} 
\end{figure}

\section{Closed-Loop Solution - Dynamic Programing} 
We now detail the Dynamic Programing approach to obtaining a feedback-optimal
control. What this amounts to is that the controller can be continuously updated
depending on the realization of the stochastic process, $X_t$.

Given a time $t$ and a realized voltage value $X_t = x$, let $\trem$ be the
(unknown) remaining time to spike, $\trem = \ts - t$. 
Starting from an arbitrary $t, x$, our remaining-cost
objective, $J[\a(\cdot); x,t]$, will be
\begin{equation}
J[\a(\cdot); x,t]  = 
\Exp\left[
\big(\trem - (\T-t) \big)^2 
+
\e \int_t^\ts  \a^2(s) \intd{s}
\,\Big|\, X_t = x
\right] 
\label{eq:OC_LS_cost_to_go}
\end{equation}
i.e.\ if time $t$ has elapsed without a spike, we now want to minimize
the difference between $\t = \ts - t$ and ($\T-t$), given the current state $x$.
Recall that in the closed-loop scenario, we assume that the value of $x = X_t$
is known to the controller.

\subsection{Hamilton-Jacobi-Bellman equation}
The Hamilton-Jacobi-Bellman (HJB) equation,
see \cite{Fleming1975,Evansb,Krylov2008}, \{ALEX: Should we also cite
Moehlis' papers that use/discuss HJB \cite{Nabi2013a,Danzl2009}???\}
associated with \cref{eq:OC_LS_cost_to_go} is obtained as follows. We introduce the value function, $\v(x,t)$
% \footnote{Traditionally, the value function is denoted as $v$. We
% use $\v$ since we are working in a neural context, where $v$ is so often used
% to denote a voltage}
:
\begin{align}
\v(x,t) =&
 \min_{\a(\cdot)_{s \geq t}}
 \{J[\a(\cdot); x,t] \}
\notag
\\
=&
\min_{\a(\cdot)_{s \geq t}}
\Exp\left[
\big(\trem - (\T-t) \big)^2 
+
\e \int_t^\ts  \a^2(s) \intd{s}
\,\Big|\,X_t = x
\right]
\label{eq:OC_LS_mean_value}
\end{align}

Then $\v$ satisfies the following HJB partial differential equation (PDE) (see
\cite{Fleming1975,Evansb,Krylov2008} for details):
\begin{equation}
\begin{gathered}
\di_t \v(x,t) + \tfrac{\b^2}{2} \di_x^2 \v(x,t) + \\
\min_{\a(x,t) \in [\amin,
\amax]}\Big\{ \e \a^2 (x,t)+ \left(\m + \a(x,t)-\tfrac{x}{\tc}\right) \di_x
\v(x,t) \Big\} = 0
\label{eq:OC_LS_mean_HJB}
\end{gathered}
\end{equation}
The special feature of \cref{eq:OC_LS_mean_HJB} in contrast to a generic
parabolic PDE is that it contains an embedded optimisation that depends on the
solution, $\v$. That is for each $x,t$ in the computational domain,
$\a$ is chosen as to minimize $\{\e \a^2 + \left(\m + \a - \tfrac{x}{\tc}\right) \di_x \v \}$. 
Here the we can solve for the optimal control, $\a^*(x,t)$ analytically as:
\begin{align}
\a^*(x,t)  =& \argmin_{\a \in [\amin, \amax]}\big\{\e \a^2 + (\m + \a-\tfrac{x}{\tc}) \di_x\v\big\}
\notag
\\
=&
\min \left(\amax, \max\left(\amin, -\frac{\di_x\v(x,t)}{2\e}\right)\right)
\label{eq:OC_LS_variance_energy_quadratic_control}
\end{align}
With this the HJB PDE becomes
\begin{equation}
\di_t \v(x,t) + \tfrac{\b^2}{2} \di_x^2\v+
\e (\astar(x,t))^2 + (\m + \astar(x,t)-\tfrac{x}{\tc}) \di_x\v
= 0
\label{eq:OC_LS_mean_HJB_bounded_control}
\end{equation}

Now, we need to consider boundary conditions (BCs) for $\v$. If $X_t = \xth$
then we have a spike now and we know that $\ts = t$. Thus
$$ \v(\xth,t) =
(t-\T)^2.$$ 
At the threshold, the value function equals the squared difference
between the desired spike time and the realized one.
Now for $x \ll .0$, we can assume that $\v$ is not
significantly affected by the change in $x$, i.e.\ that $$ \di_x \v(\xmin, t) =
0 $$ for some lower boundary $\xmin$. Such a boundary condition will be
justified if we impose artificial reflecting boundaries for the diffusion at $\xmin$. So we need to
determine for which $\xmin$ is a reflecting boundary consistent with the
dynamics. For example, we can take $\xmin$ to be two standard deviations
below the mean of the stationary distribution of the maximally
inhibited, $\a=\amin$, \cref{eq:current_mu_alpha}. That is we set $\xmin =
\tc(\mu + \amin) - 2 \b / \sqrt{\tc/2}$. In fact, we further enforce that
$\xmin \leq -0.5$.
% Ooops, we need $\di_x\v\neq 0$ and here we are setting $\di_x\v= 0$ in one of
% the BCs\ldots oh well.

Now we come to the Terminal Conditions (TCs) for $\v$. The idea is simple:
if we reach $\T$ without having spiked we apply maximum control in the
positive direction, i.e.\ $$t>T \implies \a(t) = \amax$$ Thus: 
\begin{equation}\v(x,\T) =
\Exp \Big[\trem^2 \,\Big|\, X_\T = x, \a(t) = \amax \Big].
\label{eq:valuefun_TCs}
\end{equation} 
Note that we are making an approximation here
- that is we are ignoring the energy term, $\eps u^2$ in the objective for
$t>T$. Naturally, this approximation is ever more accurate for $\eps \ll 1$.
We discuss in more detail the validity of the approximation in
\cref{sec:effect_of_eps}.
Alternatively, we could impose this approximate terminal condition at some $t^+
> \T$.

We will see the quantity on the right hand side of \cref{eq:valuefun_TCs}
repeatedly so we will give it a special name.
\begin{equation}
\Ttwo(x) := \Exp \Big[\trem^2 \,\Big|\, X_\T = x, \a(t) = \amax \Big].
\label{eq:Trem_squared}
\end{equation} 
$\Ttwo$ is the second moment of the remaining time to reach the
threshold starting at $X_\t = x$ and applying $\amax$ throughout. This quantity can be found
easily for all $x$ in the domain by solving a stationary backward Kolmogorov
equation. We show the details in \cref{sec:valuefun_TCs}.

Thus, we restate the HJB equation in its fully specified form:
\begin{equation}
\begin{gathered}
\di_t \v(x,t) + \tfrac{\b^2}{2} \di_x^2\v(x,t)+
\e \a^2(x,t) + \big(\m + \a(x,t) -\tfrac{x}{\tc}\big)\cdot \di_x\v(x,t)
= 0
\\
\a (x,t) = \min \left(\amax, \max\left(\amin, -\frac{\di_x \v(x,t)}{2\e}\right)
\right)
\\
\begin{cases}
\v(\xth,t) = (t-\T)^2  \quad & \textrm{upper BC}
\\
\di_x \v(\xmin, t)  = 0  \quad &\textrm{lower BC}
\\
\v(x,\T)  =
% \Exp \Big[\trem^2 \,\Big|\, X_\T = x, \a(t) = \amax \Big] 
% \Exp[\t^2|x, \amax]  \quad
\Ttwo(x)
\,& \textrm{TC}
\end{cases}
\end{gathered}
\label{eq:OC_LS_HJB_full}
\end{equation}
We are solving $\v(x,t)$ over the domain $[\xmin, \xth] \times [0,T]$.

\subsection{The numerical method for the HJB equation}
We now have a PDE for $\v$ and an algorithm for computing all the BCs and TCs of
this PDE. It is time to discuss the numerical method for solving \cref{eq:OC_LS_HJB_full}. Since we are in 1-d the simplest thing to do is try a
centred Finite Difference scheme using Crank-Nicholson to step in time.
Although, it might become necessary, we do not upwind the first order
derivatives for now. Also to resolve the non-linearity, $(\di_x\v)^2$, in the
PDE, we treat it as a mixed implicit-explicit term
$$(\di_x \v(x,t_k))^2 = \di_x
\v(x,t_k) \cdot \di_x \v(x,t_k) \approx \underbrace{\di_x \v(x,t_k)}_{\textrm{implicit}}
\cdot \underbrace{\di_x\v(x,t_{k+1})}_{\textrm{explicit}}.$$ Note that the
implicit term is in the previous time $t_k$ instead of, as is conventional,
the next $t_{k+1}$, because we are solving for $\v$ {\sl backwards} in time. 
TODO: Give justification for this trick.

The numerical scheme is implemented in Python, using the Scipy/Numpy library,
\cite{Scipy}.

TODO(ALEX): Mention time- space- discretization

We next demonstrate solutions for the value function and the associated
control law for a parameter set from each regime in \cref{tab:regimes}.

\subsection{Solutions of the HJB equation}
We set $\T=1.5$ and $\eps = .001$, so that the primary focus is the accurate
spiking and energy minimization considerations are secondary.

In \cref{fig:HJB_4regimes_value_control_cuts}, we depict three snapshots
in time of the value function and the control, $\v$ and $\a$, at $t = [0,\T/2,
\T]$, i.e.\ at the beginning, middle and end of the relevant time interval. Then in
\cref{fig:HJB_4regimes_value_surf} we show the entire value function, $\v(x,t)$,
the solution to \cref{eq:OC_LS_HJB_full}, for each of the parameter regimes.
% THE CUTS:
\begin{figure}[htp]
\begin{center}
  \includegraphics[width=.99\textwidth]{Figs/HJB/Regimes_vc_cuts.pdf}
  \caption[labelInTOC]{Snapshots of the value function, $\v$, on the left and on
  the optimal control, $\a$, on the right for $t$ fixed, at the start (blue),
  mid-point (green) and end-point (red), for each of the four regimes. The desired spike time is set to $\T=1.5$, the energy penalty, $\eps
  = .001$. The control bounds are $\a \in [-2,2]$. Note that the green and blue
  curves in E) are lying on top of each other for all $x$.
  \\
  A,E) Supra-Threshold-High-Noise B,F) Supra-Threshold-Low-Noise 
  C,G) Sub-Threshold-High-Noise D,H) Sub-Threshold-Low-Noise
  }   
\label{fig:HJB_4regimes_value_control_cuts}
\end{center}
\end{figure}  
% THE SURFS:    
\begin{figure}[h!]  
\begin{center}
\includegraphics[width=0.99\textwidth]{Figs/HJB/Regimes_valuesurf.pdf}
\caption{The numerical solution for the value function $\v(x,t)$ HJB PDE,
\cref{eq:OC_LS_HJB_full}, for the four different parameter regimes. Recall that
the solution is solved backwards in time.
The desired spike time is set to $\T=1.5$.  
The control bounds are $\a \in [-2,2]$. 
The contours for fixed $t$ plotted in thick correspond to the graphs in \cref{fig:HJB_4regimes_value_control_cuts}.
\\   
A) Supra-Threshold-High-Noise B) Supra-Threshold-Low-Noise 
C) Sub-Threshold-High-Noise   D) Sub-Threshold-Low-Noise  }
\label{fig:HJB_4regimes_value_surf}
\end{center} 
\end{figure}

Let us discuss the most salient features in
\cref{fig:HJB_4regimes_value_surf,fig:HJB_4regimes_value_control_cuts}. Recall
that lower values for the value function, $\v(x,t)$, are preferred, since we are
minimizing. We see the same general shape in all four space-time surface plots
of the value function in \cref{fig:HJB_4regimes_value_surf}. At the end of the
interval, $\v(x,\T)$ is monotonically decreasing in $x$, which is to be expected
given its terminal condition in \cref{eq:valuefun_TCs}. That is the lower the
value of $X_\T$ the more on average will we have to wait for the spike to occur.
% This same basic shape represents the idea that we would like to be close to
% the threshold, $x = 1$, near the end of the interval when $t \approx \T$, but
% we would like to be far from threshold, for $t \ll \T$. Indeed, at the very
% end of the interval, when $t = \T$, the value function, $v(x,\T)$, is
% monotonically decreasing in $x$, because the lower the value of $X_\T$ the
% more time on average will it take us to spike and the more delay will we have
% from our desired spike time.
As we go back in time $\v$ inverts, with a clear peak near the upper, threshold
boundary. That is, for intermediate values of $t$ we have to consider the risk
of spiking too early, represented by the high value of $\v$ near the upper
boundary {\sl and} the risk of spiking too late, represented by the high values
of $\v$ at the lower boundary. As we progress even further back in time to the
beginning of the interval, the peak near the threshold rises further, since we are spiking
{\sl earlier}, while the peak at the lower end flattens, since now there is
enough time to reach threshold despite starting far from it. The full surface
plots for $\v(x,t)$ in \cref{fig:HJB_4regimes_value_surf} can be thought of as
the interpolation between these three basic phases.

% Note, that exact shape is generated by Since we have put a very small weight
% on the energy penalty, $\e = .001$, and we have a significantly high bound on
% the maximum / minimum of the applied control, the value function becomes quite
% flat for $x \ll 1$ and $t \ll \T$, since it is likely that we can apply enough
% control to avoid an early spike and still drive the system close to threshold
% at the desired time, $t = \T$
Now focus on the controls, on the right in
\cref{fig:HJB_4regimes_value_control_cuts}. Naturally, at the end of the
interval, the control takes its maximum value, $\a(x,\T) = \amax$ for all $x$,
i.e.\ it gives the maximum available push for the neuron to spike. This is
implied \cref{eq:valuefun_TCs}. As we go back in time, however, the control $\a$
decreases for $x \ll \xth$, and it becomes negative, i.e.\ inhibitory for $x
\lesssim \xth$. That is intuitively consistent with the problem objective - for
$t<\T$, we want to bring $X_t$ closer to the threshold, but not too close, to
avoid early spiking!

Now let us consider the differences between the individual regimes, i.e.\ the
effect of changing the bias and the noise intensity, $\m,\b$. All things being
equal, the effect of increasing the noise intensity, $\b$, is to lift the value
function, i.e.\ to make our objective worse, at the beginning of the interval
and to decrease it at the end. This is to be expected since we are attempting
to minimize the variance in the spike time and without noise there would be no
variance at all. However, at the end of the interval, noise only helps with the
spiking, since now we are only interested in spiking as soon as possible, and a
higher noise intensity, will tend to decrease the spike time on average.
Increasing $\b$ also has the effect of increasing the size of the boundary layer
near $\xth$, where the value function rises steeply. I.e.\ for small $\b$ that
layer is small since the risk of spiking early is only significant close to
the threshold. For larger $\b$ this layer increases, see panels A,C, with a thin
layer, vs.\ B,D, with a thicker layer, in
\cref{fig:HJB_4regimes_value_control_cuts}. Similarly, increasing the bias,
$\mu$, tends to decrease the value function, especially at the end since that
has the unequivocal effect of preventing late spikes.
\clearpage

\section{Open-loop Stochastic Control}
When the value of $X_t$ is unobservable, the best
one can do is to use the transition density to perform the optimization. Since the
transition density follows a (deterministic) PDE, we apply a Maximum
Principle for PDEs as a method of obtaining the optimal control,\cite{Borzi2012}.

\subsection{Fokker-Planck equation for the state density evolution}
We write the transition density of $X$, conditional on no
spikes having occurred as:
\begin{align*}
f(x,t) \intd{x} &= \Prob[X_t\in \intd{x} \,\big|\, X_0 = 0,\, X_{s} < \xth \quad
\forall s < t]
\end{align*}
Then $f$ satisfies a Fokker-Planck equation with absorbing boundaries,
see Ch.\ 7 in \cite{Jacobs},
\begin{equation}
\begin{gathered}
\di_t f(x,t) =
				\frac{\b^2 }{2}\cdot \di^2_x f -
				\di_x \Big[ \left( \m + \a(t)- \frac{x}{\tc}\right)  \cdot f \Big]
\\
\\
\begin{array}{lll}
	&\textrm{st.\ }&
	\left\{ \begin{array}{lcl}
	 f(x,0) &=& \delta(x) \quad \textrm{delta function}
	\\
	(\m + \a(t)- \tfrac x \tc)f - \di_x \tfrac {\b^2}2 f] \big|_{x=\xmin} &\equiv&
	0 \quad \textrm{reflecting BCs at some } \xmin
	\\
	f(x,t) |_{x=\xth} &\equiv& 0 \quad \textrm{absorbing BCs at } \xth
\end{array} \right.
\end{array}
\label{eq:FP_pde_OU_PDF}
\end{gathered}
\end{equation}
In theory, $\xmin = -\infty$, but in the numerics below we will need to
truncate it to some finite value, exactly as in the HJB equation,
\cref{eq:OC_LS_HJB_full}.
Note that we write $\a(t)$ here instead of $\a(x,t)$ since we cannot use the
value of $X_t$ in choosing the optimal control.

The $f$ dynamics can also be written as
$$
\di_t f(x,t) = - \di_x \phi(x,t)
$$
for the probability flow
$$
\phi(x,t) = (\m + \a(t) - \tfrac x \tc)f - \di_x [\tfrac {\b^2}2 f]
$$
Then the lower BC is
$$
\phi(x,t) |_{x=\xmin} \equiv 0
$$

We will also need a short hand notation for the differential operator on the
right side of \cref{eq:FP_pde_OU_PDF}. Let
$$ \L_\a[\cdot] := \di^2_x \Big[\frac{\b^2 }{2} \cdot\Big] -
 \di_x \Big[ (\m + \a(t)- \frac{x}{\tc}) \cdot \Big] $$where
 $[\cdot]$ indicates the argument of the operator $\L_\a$. We will usually
 omit the subscript $\a$, but have written it now to emphasize that the
 differential operator is parametrized by the control, $\a$.

\subsection{Restating the objective in terms of the transition density}
Let us recall our objective, \cref{eq:OC_LS_variance}: 
$$
J[\a(\cdot)] = \Exp\left[
(\ts - \T \big)^2
+
\e \int_0^\ts  \a^2(s) \intd{s}
\right].
$$
Rewriting this in terms of the transition density, $f$, reads:
\begin{align}
J[\a(\cdot)] =&
\int_\xmin^\xth \Ttwo(x) \cdot f(x,\T) \intd{x}
\notag
\\
&+ \int_0^\T \phi(\xth, t) (t-\T)^2 \intd{t}
\label{eq:OC_LS_variance_density}
\\
&+  \e \int_0^\ts  \a^2(t)  \int_\xmin^\xth f(x,t) \intd{x} \intd{t}
\notag
\end{align}
Let us explain in more detail each term on the RHS in
\cref{eq:OC_LS_variance_density}.

The first term $$ \int_\xmin^\xth \Ttwo(x) \cdot f(x,\T)\intd{x} $$ counts the cost of trajectories which spike too late. This cost is the expected
squared time-to-hit starting at $x$, with  $\a = \amax$, weighted by the
probability of $X_\T = x$, which is just $f(x,\T)$. Recall that $\Ttwo$ is
defined in \cref{eq:Trem_squared}.

The second term $$ \int_0^\T \phi(\xth,t) (t-\T)^2 \intd{t} $$ counts the cost
of trajectories which spike too early, that is the squared difference between
some realized spike time, $\ts = t$, and desired spike time $\T$, weighted by
the probability of a spike at $t$ which is just the outflow of probability,
$\phi(\xth,t)$. Recall that we assume that $\xth = 1$ throughout. Note further
that due to the homogeneous Dirichlet BC at $\xth$, $f(\xth) = 0$, the outflow
is simply: $$ \phi(\xth, t) = -\frac{\b^2 }{2} \di_x f(\xth, t).$$

Finally, the third term
$$
\e \int_0^\ts  \a^2(t)  \left[  \int_\xmin^\xth f(x,t) \intd{x} \right] \intd{t}
$$
is the energy cost: the inner integral, $\int_\xmin^\xth f(x,t) \intd{x}$
takes into account that we incur an energy cost only for those trajectories
which have not yet spiked.

With that our optimal control $\a^*(\cdot)$ will naturally be found via:
$$
\a^*(\cdot) = \argmin_{\a(\cdot)} J[\a(\cdot)]
$$

\subsection{Optimizing using a Maximum Principle}
\label{sec:PDE_max_principle_for_pdf}
By now our stochastic optimal control problem modelled by SDEs
has become a deterministic optimal control problem modelled by PDEs. Our control
$\a$ influences the evolution density $f$ and via $f$, the integrals in the
objective, $J$.

The Maximum Principle for PDEs, which is an extension to the famous Pontryagin
Maximum Principle from finite dimensional systems, introduces an adjoint
variable, $p$, which solves a PDE related to the PDE satisfied by the density
$f$ and then calculates the optimal control, $\a(\cdot)$, as a functional of $f$
and $p$.

In short, the equation for the adjoint function, $p$, is
\begin{equation}
\begin{gathered}
\begin{aligned}
\di_t p =& - \Lstar[p]
\\
		=&
			- \Big[ \frac{\b^2 }{2}\cdot \di^2_x p +
			(\m + \a(t)- \frac{x}{\tc})  \cdot \di_x p \Big]
\end{aligned}
\\
\st
\begin{cases}
	p(x,\T) &= \Ttwo(x)
	\\
	\di_x p  \big|_{x=\xmin} &\equiv 0
	\\
	p \big|_{x=\xth} &\equiv (s-T)^2
\end{cases}
\label{eq:adjoint_pde_OU}
\end{gathered}
\end{equation}

and then $\a$ can be found via
\begin{equation}
\Big\{
 \tilde{\e}  2 \a(t)
+ p f \Big|_\xmin
- \int _\xmin^\xth p \cdot \di_x f \intd{x}
\Big\} = 0
\quad \forall t \in [0,T]
\label{eq:J_necessary_condition}
\end{equation}

More practically, the quantity,
$$\di_\a J =  \Big\{
 \tilde{\e}  2 \a(t)
+ p(x,t) f(x,t) \Big|_\xmin
- \int _\xmin^\xth p \cdot \di_x f(x,t) \intd{x}
\Big\}
$$
gives the direction of increase of $J$ at $\a(t)$
and we can use it as a gradient in a descent algorithm, given some initial
guess, $\a_0(t)$, for the control.

Finally, we are ready to calculate the open-loop stochastic optimal control for
the four parameter regimes. The calculated optimal controls for each regime are
in \cref{fig:FBK_Regimes_cs}. Recall that the optimal control, $\a^*(t)$, is
open-loop and thus it is only a function of time.
% \usepackage{graphics} is needed for \includegraphics 
\begin{figure}[htp]
\begin{center}
  \includegraphics[width=.99\textwidth]{Figs/FP_Adjoint/Regimes_cs.pdf}  
  \caption[labelInTOC]{The deterministic optimal controls for each parameter
  regime as functions of time, $t \in [0, \T]$.
  The desired spike time is set to $\T=1.5$, the energy penalty, $\eps
  = .001$ and the bounds are $\a \in [-2,2]$.
  A) Supra-Threshold-High-Noise B) Supra-Threshold-Low-Noise 
C) Sub-Threshold-High-Noise   D) Sub-Threshold-Low-Noise  }
  \label{fig:FBK_Regimes_cs}   
\end{center}
\end{figure}

Let us discuss the most salient features of the controls calculated in
\cref{fig:FBK_Regimes_cs}. In general, the strategy is to inhibit the neuron at
the beginning of the interval, $\a(t) = \amin$ for $t \ll \T$ and to excite it
near the end, $\a(t) = \amax$ for $t \approx \T$. The smooth portion that connects these
two segments in the middle of the interval is due to the energy penalty, $\e \int
\a^2(s) \intd{s}$. The behaviour of the control in different regimes
is relatively simple to explain and we see what we would expect. In a
sense the bias current $\mu$ can be absorbed by the control and so increasing
$\mu$ amounts to decreasing $\a$ modulo its bounds. That is exactly the
difference between plots A) vs.\ C) and B) vs.\ D) in \cref{fig:FBK_Regimes_cs},
where the reduction of $\mu$ in the lower plots, C) and D), results in an
increase in $\a(t)$ compared to A) and B), respectively. The effect of the noise
intensity, $\b$, is more subtle and it is not the same in the Supra vs.\ the
Sub-Threshold regimes A,B) vs.\ C,D). In the Supra-Threshold regime,  panels
A,B), reducing $\b$, reduces the need for excitatory  control towards the end,
since the bias alone should put the neuron over the threshold. Thus the main
part of the control with low noise in the Supra-Threshold regime is to stop the
neuron from spiking too early. In the Sub-Threshold regime, panels C,D), the
situation is somewhat reversed - reducing $\b$, obviates the need to apply an
inhibitory control in the early part of the interval and also prompts the
controller to apply excitatory input earlier, since it is the controller which
is now the main drive for a spike.

% An example for . The results are in
% \cref{fig:FP_adjoint_objective_control_convergence}.
% \begin{figure}[h]
% \begin{center}
% \subfloat[$\a_k(t)$ ]
% {
% \includegraphics[width=0.48\textwidth]
% {Figs/FP_Adjoint/ExampleControlConvergence_control.png}
% }
% \subfloat[$J_k$ ]
% {
% \includegraphics[width=0.48\textwidth]
% {Figs/FP_Adjoint/ExampleControlConvergence_objective.png}
% }
% \caption[ ]{The result of an entire optimization iteration, on the left the
% iterates of the control, $\a_k(t)$, on the right the progress of the
% objective with each iteration. The final control, in purple, is inhibitory at
% the beginning of the time interval, $\a < 0$, unlike the the deterministic
% control solution.
% We see a significant
% reduction in the objective value, $J$, and thus an improvement, on the
% order of 50\%.}
% \label{fig:FP_adjoint_objective_control_convergence}
% \end{center}
% \end{figure}
% 
% \begin{table}[h]
% \centering
% \begin{tabular}{lc}
% Control Law & Squared Error \\
% \hline
% Deterministic &  0.647 \\
% Stochastic &  0.336\\
% Theory (Value function) & 0.285
% \end{tabular}
% \caption{Realized performance of the different control laws and the theoretical
% expected performance of the stochastic law (last row)}
% \label{tab:realized_avg_errors_det_vs_stoch}
% \end{table}
% 
% \begin{figure}[h]
% \begin{center}
% \subfloat[A]
% {
% \label{fig:controlled_traj_ex1}
% \includegraphics[width=0.33\textwidth]
% {Figs/ControlSimulator/example_controlled_trajectories_id1.png}
% }
% \subfloat[B]
% {
% \label{fig:controlled_traj_ex2}
% \includegraphics[width=0.33\textwidth]
% {Figs/ControlSimulator/example_controlled_trajectories_id4.png}
% }
% \subfloat[C]
% {
% \label{fig:controlled_traj_ex3}
% \includegraphics[width=0.33\textwidth]
% {Figs/ControlSimulator/example_controlled_trajectories_id3.png}
% }
% \caption[]{Examples for the controlled trajectories using both the deterministic
% and the stochastic control approaches. The red vertical line in the plots
% indicates the desired spike-time, $\T$. In A, B the performance of both
% control laws is essentially the same, but in C we see the advantage of the
% stochastic approach. $\tc, \b = [.75, 1.25]$.}
% \label{fig:control_trajectories_examples}
% \end{center}
% \end{figure}
% \begin{figure}[htp]
% \begin{center}
%   \includegraphics[width=.9\textwidth]{Figs/ControlSimulator/example_controlled_trajectories_hists.png}
%   \caption[labelInTOC]{Histogram of the spike timing error for the
%   deterministic (left) vs. the stochastic (right) control laws.}
%   \label{fig:error_histograms_det_vs_stoch}
% \end{center}
% \end{figure}

\clearpage
\section{Effect of energy penalty}
\label{sec:effect_of_eps}
So far we have assumed a low value for the energy penalty parameter,
$\eps$, in the objective defined in \cref{eq:OC_LS_variance_energy}. Our
approach has been that we are most concerned with the accurate spiking and that
penalizing energy is rather intended to regularize the control and avoid
excessive chattering of the control between its extreme values, than as a goal
in minimizing energy in its own right. Now, we take some time to explore the
effect of a higher $\eps$ on the optimal controls. Intuitively, a
higher value of $\eps$ will tend to bring the optimal control, $\a^*$, closer
to zero. We make a comparison for the four regimes for both the
closed-loop control and the open-loop control in
\cref{fig:HJB_4regimes_control_different_eps,fig:FBK_Regimes_cs_different_es}.
Indeed, for both the closed-loop and open-loop optimal controls, increasing
$\eps$ tends to reduce the absolute value of $\a^*(t)$. In particular, for the
closed-loop control it tends to broaden the area of transition for
decreasing $x$ when the control swings from its minimal, i.e.\ inhibitory value,
to its maximal, i.e.\ excitatory value. Similarly for the open-loop control,
instead of banging from its minimal bound at the beginning of the interval to
its maximal bound at the end, the optimal control for $\e = .1$ tends to have a
more mild transition from a slightly inhibitory to slightly excitatory
values.

Note however that increasing $\eps$ has an important effect on the validity of
the approximation, used to form the terminal conditions for
the value function and the adjoint variable, $\v,p$. We assumed that applying
$\amax$ is optimal for $t>\T$. This is indeed true in the
limit $\eps \ra 0$. However, even for a finite value of $\eps$, this may still
be the optimal thing to do. Indeed, in the original calculations,
where $\eps=.001$, see the panels on the left in
\cref{fig:HJB_4regimes_control_different_eps,fig:FBK_Regimes_cs_different_es},
the so-obtained value function {\sl implied} that $\a(x,\T) = \amax.$. I.e.\ our
guess that $\a(x, t) = \amax$ for $t > \t$, is self-consistent with the
so-obtained value function. In a sense, this is like confirming an ansatz. For a higher
value of $\eps$, like $\eps = .1$, this is no longer the case. The red
curves on the right-side of \cref{fig:HJB_4regimes_control_different_eps} are no
longer pushed up at $\a = \amax$, and as such it is no longer valid to
assume that the value function can be obtained by assuming $\a(t>\T) = \amax$
and then just calculating the expected remaining time-to-spike.
The exact same conclusion can be drawn from the open-loop control. There the
optimal control always (almost) reaches its maximum, $\amax$ while choosing
$\eps = .001$, but no longer so for the higher $\eps = .1$.
In fact, as we mentioned after \cref{eq:valuefun_TCs}, the correct thing to do
for higher values of $\eps$, is to push the
calculation interval to some $t^+ > \T$, apply the same terminal condition at
$t^+$ instead of $\T$ and then solve backwards (either for the closed- or
open-loop control). One will be guaranteed to find some $t_f > \T$ that works
since eventually the quadratic term $(t-\T)^2$ will dominate the energy term in
the objective, which is linear in $t$. We do not explore this further here. 
% CLOSED_LOOP EPS: 
\begin{figure}[htp]
\begin{center}
  \includegraphics[width=.99\textwidth]{Figs/HJB/Regimes_eps_comparison.pdf}
  \caption[labelInTOC]{Effect of $\e$ on the closed-loop control. Snapshots of
  $\a(x,t)$ for $t$ fixed, at the start (blue), mid-point (green) and end-point
  (red), for each of the four regimes and different values of $\eps$.
  On the left, we use $\eps = .001$ and on the right $\eps = .1$. 
  The desired spike time is set to
  $\T=1.5$ and the bounds are $\a \in [-2,2]$.
  A,B) Supra-Threshold-High-Noise C,D) Supra-Threshold-Low-Noise 
  E,F) Sub-Threshold-High-Noise G,H) Sub-Threshold-Low-Noise}
\label{fig:HJB_4regimes_control_different_eps} 
\end{center}
\end{figure}
% OPEN_LOOP EPS:
\begin{figure}[htp]
\begin{center}
  \includegraphics[width=.99\textwidth]{Figs/FP_Adjoint/Regimes_eps_comparison.pdf}
  \caption[labelInTOC]{Effect of $\e$ on the open-loop control. 
  The deterministic time-trajectory of $\a(t)$ for each
  parameter regime as function of time, $t \in [0, \T]$.
  On the left, we use $\eps = .001$ and on the right $\eps = .1$.
  The desired spike time is set to $\T=1.5$ and the bounds are $\a \in [-2,2]$.
  A,B) Supra-Threshold-High-Noise C,D) Supra-Threshold-Low-Noise 
  E,F) Sub-Threshold-High-Noise G,H) Sub-Threshold-Low-Noise}
  \label{fig:FBK_Regimes_cs_different_es}
\end{center}
\end{figure}

\clearpage

\section{Test Comparison of Different Controls}
\label{sec:probabilistic_numerical_test}
Having obtained the optimal controllers, both closed- and open-loop, we evaluate
their performance with simulated realizations of the voltage process,
\cref{eq:X_evolution_uo}. In particular, while they both minimize the expected
squared deviation of the spike-time from some desired spike time, we analyze
the distribution of the squared deviation and the behaviour of the
controls for different parameter regimes.

In addition to the optimal controllers, we also show the behaviour of
another control law - perhaps the most naive one - which is obtained by ignoring
the noise and the energy penalty. That is we find a {\sl constant} value of $\a$
which satisfies the desired boundary conditions, $x(0) = .0, x(\T) = \xth$. This naive
controller will be called 'Deterministic', since it assumes deterministic
dynamics in $X_t$, i.e.\ $\b = 0$.

For the comparison, we sample $N$ realizations of the controlled system
and apply in turn each of the three controls. Naturally, we reuse the same
realization of the underlying stochastic process, $W_t$, for each of the
three different controls.

We set $\e = .001$ and $\amax = 2.0$. As such the energy cost is of secondary
importance and the paramount effect on the objective is the difference $(\ts -
\T)$, which we are trying to minimize.

The performance of the three controllers for each parameter regime is given in
\cref{tab:realized_avg_errors_det_vs_openloop_vs_stoch} and  
\cref{fig:error_histograms_det_vs_openloop_vs_stoch}. Naturally, the closed-loop
achieves a lower error than the open-loop, and the naive controller fares worst.
The difference in performance mostly depends on the strength of the noise. Thus,
for low noise, the performance of the stochastic controllers, be they open- or
closed-loop is not much superior to the naive deterministic controller. On the
contrary - in the high-noise regime, using a stochastic controller gives a much
lower error on average between the desired $\T$ and the realized $\ts$.

For illustration sake, we also visualize some trajectories from the
Sub-Threshold-High-Noise regime in \cref{fig:control_trajectories_examples}.

\begin{table}[h]
\begin{center}
\subfloat[Supra-Threshold High-Noise]{
\input{Figs/ControlSimulator/MeanSquaredErrors__SUPT_HN.txt}}\\
\subfloat[Supra-Threshold Low-Noise]{
   \input{Figs/ControlSimulator/MeanSquaredErrors__SUPT_ln.txt}}\\
 \subfloat[Sub-Threshold High-Noise]{
 \input{Figs/ControlSimulator/MeanSquaredErrors__subt_HN.txt}}\\
 \subfloat[Sub-Threshold Low-Noise]{
 \input{Figs/ControlSimulator/MeanSquaredErrors__subt_ln.txt}}\\
%  \begin{tabular}{lcc}
% Control Law &  Squared Error &  Squared Error \\
%  & Empirical & Theoretical\\% \hline
% Deterministic &  0.621 & - \\
% Open-Loop Stochastic & 0.356 & 0.382\\
% Feedback Stochastic &  0.288& 0.285\\
% \hline % \end{tabular} -
\caption{TODO(ALEX): THIS TABLE IS MOSTLY A SANITY CHECK TO SEE THAT THE CODE
WORKS - CONSIDER OMITTTING IN DRAFT.
Realized and theoretical performance of the different control laws. The
empirical performance is obtained using $N=10000$ sample paths. The theoretical
performance is found using the optimal value for $J$ for the open-loop
stochastic control and the value function $\v(x=0, t =0)$ for the closed-loop
stochastic control.}
\label{tab:realized_avg_errors_det_vs_openloop_vs_stoch}  
\end{center}
\end{table}

\begin{figure}[h]
\begin{center}
\subfloat[Supra-Threshold High-Noise]{
  \includegraphics[width=.99\textwidth]
  {Figs/ControlSimulator/Regimes_SUPT_HN_errors_hist.pdf}
  }\\
\subfloat[Supra-Threshold Low-Noise]{
  \includegraphics[width=.99\textwidth] 
  {Figs/ControlSimulator/Regimes_SUPT_ln_errors_hist.pdf}
}\\
\subfloat[Sub-Threshold High-Noise]{
  \includegraphics[width=.99\textwidth]
  {Figs/ControlSimulator/Regimes_subt_HN_errors_hist.pdf}
}\\   
\subfloat[Sub-Threshold Low-Noise]{
  \includegraphics[width=.99\textwidth]
  {Figs/ControlSimulator/Regimes_subt_ln_errors_hist.pdf}
}
  \caption[labelInTOC]{Histogram of the spike timing error for the
  deterministic (left) vs. open-loop stochastic(centre) vs. closed-loop
  stochastic (right) control laws. This is for the same problem as in
  \cref{fig:control_trajectories_examples}. 
  We have used $N=10000$ sample paths
  to form the statistics}
  \label{fig:error_histograms_det_vs_openloop_vs_stoch}
\end{center}
\end{figure}

\begin{figure}[h]
\begin{center}
\subfloat[]
{
\label{fig:controlled_traj_ex1}
\includegraphics[width=0.33\textwidth]
{Figs/ControlSimulator/SubTHighNoise_Traj5.pdf}
}
\subfloat[ ]
{
\label{fig:controlled_traj_ex2}
\includegraphics[width=0.33\textwidth] 
{Figs/ControlSimulator/SubTHighNoise_Traj3.pdf}
}
\subfloat[ ]
{
\label{fig:controlled_traj_ex3}
\includegraphics[width=0.33\textwidth]
{Figs/ControlSimulator/SubTHighNoise_Traj7.pdf}
}
\caption[]{Examples for the controlled trajectories using the
deterministic, open-loop stochastic and closed-loop stochastic control
approaches. The black vertical line in the plots indicates the desired spike-time, $\T$.
For the parameter values are $\m, \tc, \b = [.2, .5,  1.5]$ (Sub-Threshold
High-Noise regime). $\T = 1.5, \eps = .001$. The three panels from left to
right are three different realizations of the model dynamics. On the upper
plots, we show the voltage evolution, $X_t$, on the lower plots we show the
applied control, $\a(t)$. Note that $\a(t)$ obtained from the
deterministic and open-loop controls are the same for all three samples.}
\label{fig:control_trajectories_examples} 
\end{center}
\end{figure}

\begin{figure}[h]
\begin{center}
\subfloat[]
{
\label{fig:controlled_traj_ex1}
\includegraphics[width=0.33\textwidth]
{Figs/ControlSimulator/HighEps_Traj5.pdf}
}
\subfloat[ ]
{
\label{fig:controlled_traj_ex2}
\includegraphics[width=0.33\textwidth]
{Figs/ControlSimulator/HighEps_Traj6.pdf}
}
\subfloat[ ]
{
\label{fig:controlled_traj_ex3}
\includegraphics[width=0.33\textwidth]
{Figs/ControlSimulator/HighEps_Traj0.pdf}  
}
\caption[]{The effect of higher energy penalty, $\eps = .1$!
Examples for the controlled trajectories using the deterministic,
open-loop stochastic and closed-loop stochastic control approaches. The black vertical line in the plots indicates the desired
spike-time, $\T$. For the parameter values are $\m, \tc, \b = [.2, .5,  1.5]$ (Sub-Threshold
High-Noise regime). $\T = 1.5, \eps = .1$. The three panels from left to
right are three different realizations of the model dynamics. On the upper
plots, we show the voltage evolution, $X_t$, on the lower plots we show the
applied control, $\a(t)$. Note that $\a(t)$ obtained from the
deterministic and open-loop controls are the same for all three samples.}
\label{fig:control_trajectories_examples_high_epsilon} 
\end{center}
\end{figure}

\clearpage

\section{Control Robustness}
We have so far assumed that we know the parameters determining the dynamics in
\cref{eq:X_evolution_uo} exactly. That is we assume that the values for $\tc,
\b$ used to obtain the controls are the same as the ones that drive the
controlled dynamics. That is slightly naive, as in any realistic application
these will have to be estimated and the estimates will have some error
associated with them. 
Note that misestimating $\mu$ has the most immediate and clear effect on the
control policy as that amounts to a direct shift of the control and so for now,
let us assume that $\mu$ is estimated correctly.

Let us then discuss one of the regimes, the Sub-Threshold,
High-Noise regime, and see what is the impact of misspecifying the parameters.

We need to assume that the errors large but not too large\ldots

So say, $$\b_{est} = \b / 1.5$$
and 
$$\tc_{est} = \tc / 1.5$$

So we will have:
$$\mu, \tc_{est}, \b_{est} = \{.2, .5, 1.5 \}$$ (which is the control
calculated already for the Sub-Threshold,
High-Noise regime), while the actual simulated dynamics will be driven by
$$\mu, \tc, \b = \{.2, .75, 2.25 \}$$

We show the effect in
\cref{tab:realized_avg_errors_det_vs_openloop_vs_stoch_misspecified}.
\begin{table}[h]
\begin{center}
 \subfloat[Sub-Threshold High-Noise]{
 \input{Figs/ControlSimulator/Misspec_MeanSquaredErrors__subt_HN.txt}}\\
 \caption{The effect of misspecifying the parameters. On the left, we
 system parameters and the parameters used to obtain the control are the
 same, i.e.\ accurate, on the right, they are misspecified.}
\label{tab:realized_avg_errors_det_vs_openloop_vs_stoch_misspecified}
\end{center}   
\end{table}

\section{Trains Simulation }
We now turn to the ultimate goal of our analysis, the control of {\sl spike
trains}. In particular, we will generate $M=10$ realizations of $N=20$ spikes
each in attempt to meet a prescribed spike train. We will focus on two
regimes, the sub-threshold regime with either low or high noise. 
The results
for parameters from the Sub-Threshold, Low-Noise regime are shown in
results in \cref{fig:targettrain_cl_lownoise,fig:targettrain_ol_lownoise}.
While, the results
for parameters from the Sub-Threshold, High-Noise regime are shown in
results in \cref{fig:targettrain_cl_highnoise,fig:targettrain_ol_highnoise}.
%\usepackage{graphics} is needed for \includegraphics
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%% CRIT REGIME LOW-NOISe:
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[htp]     
\begin{center}  
  \includegraphics[width=.99\textwidth]{Figs/TrainController/CRITLN_cl_trains_sim_10.pdf}
  \caption[ ]{How good the {\sl closed-loop} controller is at hitting a target
  spike train. Uses the parameters from the CRITICAL,Low-Noise regime. The
  target spike train has ISIs that are exponentially distributed with a
  mean of $1.5$ a.u. There are $N=20$ spikes and $M=10$ realizations of the
  control strategy}
  \label{fig:targettrain_cl_critlownoise}  
\end{center}
\end{figure}    
\begin{figure}[htp]     
\begin{center}  
  \includegraphics[width=.99\textwidth]{Figs/TrainController/CRITLN_ol_trains_sim_10.pdf}
  \caption[ ]{How good the {\sl open-loop} controller is at hitting a target
  spike train. Uses the parameters from the CRITICAL,Low-Noise regime. The
  target spike train has ISIs that are exponentially distributed with a
  mean of $1.5$ a.u. There are $N=20$ spikes and $M=10$ realizations of the
  control strategy} 
  \label{fig:targettrain_ol_critlownoise}    
\end{center} 
\end{figure}   
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%% CRIT REGIME HIGH-NOISe:
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[htp]     
\begin{center}  
  \includegraphics[width=.99\textwidth]{Figs/TrainController/CRITHN_cl_trains_sim_10.pdf}
  \caption[ ]{How good the {\sl closed-loop} controller is at hitting a target
  spike train. Uses the parameters from the CRITICAL, High-Noise regime. The
  target spike train has ISIs that are exponentially distributed with a
  mean of $1.5$ a.u. There are $N=20$ spikes and $M=10$ realizations of the
  control strategy}
  \label{fig:targettrain_cl_critlownoise}  
\end{center}
\end{figure}   
\begin{figure}[htp]     
\begin{center}  
  \includegraphics[width=.99\textwidth]{Figs/TrainController/CRITHN_ol_trains_sim_10.pdf}
  \caption[ ]{How good the {\sl open-loop} controller is at hitting a target
  spike train. Uses the parameters from the CRITICAL, High-Noise regime. The
  target spike train has ISIs that are exponentially distributed with a
  mean of $1.5$ a.u. There are $N=20$ spikes and $M=10$ realizations of the
  control strategy}
  \label{fig:targettrain_ol_critlownoise}   
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%% CRIT REGIME LOW-NOISe (AHmadian):
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[htp]     
\begin{center}  
  \includegraphics[width=.99\textwidth]{Figs/TrainController/CRITLN_Ahmadian_cl_trains_sim_10.pdf}
  \caption[ ]{AHMADIAN COMPARES: How good the {\sl closed-loop} controller is at
  hitting a target spike train. Uses the parameters from the CRITICAL, High-Noise regime. The
  target spike train has ISIs that are exponentially distributed with a
  mean of $1.5$ a.u. There are $N=20$ spikes and $M=10$ realizations of the
  control strategy}
  \label{fig:targettrain_cl_critlownoise}  
\end{center}
\end{figure}   
\begin{figure}[htp]     
\begin{center}  
  \includegraphics[width=.99\textwidth]{Figs/TrainController/CRITLN_Ahmadian_ol_trains_sim_10.pdf}
  \caption[ ]{AHMADIAN COMPARES: How good the {\sl open-loop} controller is at
  hitting a target spike train. Uses the parameters from the CRITICAL, High-Noise regime. The
  target spike train has ISIs that are exponentially distributed with a
  mean of $1.5$ a.u. There are $N=20$ spikes and $M=10$ realizations of the
  control strategy}
  \label{fig:targettrain_ol_critlownoise}   
\end{center}
\end{figure}  
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%% FIGS:
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[htp]     
\begin{center}  
\subfloat[DO NOT SKIP MISSED]{
  \includegraphics[width=.99\textwidth]{Figs/TrainController/SUBTLN_cl_trains_sim_10.pdf}
  }\\
 \subfloat[SKIP MISSED]{
  \includegraphics[width=.99\textwidth]{Figs/TrainController/SUBTLN_cl_skip_trains_sim_10.pdf}
  }
  \caption[ ]{How good the {\sl closed-loop} controller is at hitting a target
  spike train. Uses the parameters from the Sub-Threshold,Low-Noise regime. The
  target spike train has ISIs that are exponentially distributed with a
  mean of $1.5$ a.u. There are $N=20$ spikes and $M=10$ realizations of the
  control strategy}
  \label{fig:targettrain_cl_lownoise}  
\end{center}
\end{figure}   
\begin{figure}[htp]  
\begin{center}
\subfloat[DO NOT SKIP MISSED]{
  \includegraphics[width=.99\textwidth]{Figs/TrainController/SUBTLN_ol_trains_sim_10.pdf}
  }\\
 \subfloat[SKIP MISSED]{
 \includegraphics[width=.99\textwidth]{Figs/TrainController/SUBTLN_ol_skip_trains_sim_10.pdf}
 }     
  \caption[ ]{How good the {\sl open-loop} controller is at hitting a target
  spike train. Uses the parameters from the Sub-Threshold,Low-Noise regime. The
  target spike train has ISIs that are exponentially distributed with a
  mean of $1.5$ a.u. There are $N=20$ spikes and $M=10$ realizations of the   
  control strategy}
  \label{fig:targettrain_ol_lownoise}
\end{center}
\end{figure}
\begin{figure}[htp]  
\begin{center}    \subfloat[DO NOT SKIP MISSED]{ 
  \includegraphics[width=.99\textwidth]{Figs/TrainController/SUBTHN_cl_trains_sim_10.pdf}  }\\
 \subfloat[SKIP MISSED]{
 \includegraphics[width=.99\textwidth]{Figs/TrainController/SUBTHN_cl_skip_trains_sim_10.pdf}
 }  
  \caption[ ]{How good the {\sl closed-loop} controller is at hitting a target
  spike train. Uses the parameters from the Sub-Threshold,High-Noise regime. The
  target spike train has ISIs that are exponentially distributed with a 
  mean of $1.5$ a.u. There are $N=20$ spikes and $M=10$ realizations of the
  control strategy}
  \label{fig:targettrain_cl_highnoise}
\end{center}
\end{figure}  
\begin{figure}[htp]  
\begin{center}    \subfloat[DO NOT SKIP MISSED]{ 
  \includegraphics[width=.99\textwidth]{Figs/TrainController/SUBTHN_ol_trains_sim_10.pdf}  }\\
 \subfloat[SKIP MISSED]{
 \includegraphics[width=.99\textwidth]{Figs/TrainController/SUBTHN_ol_skip_trains_sim_10.pdf}
 }  
  \caption[ ]{How good the {\sl open-loop} controller is at hitting a target
  spike train. Uses the parameters from the Sub-Threshold,High-Noise regime. The
  target spike train has ISIs that are exponentially distributed with a
  mean of $1.5$ a.u. There are $N=20$ spikes and $M=10$ realizations of the
  control strategy}
  \label{fig:targettrain_ol_highnoise}
\end{center}
\end{figure}    
\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%% SKIP MISSED FIGS:
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \begin{figure}[htp]     
% \begin{center}  
%   \includegraphics[width=.99\textwidth]{Figs/TrainController/SUBTLN_cl_skip_trains_sim_10.pdf}
%   \caption[ ]{SKIPPING MISSED TARGET. How good the {\sl closed-loop} controller is at hitting a target
%   spike train. Uses the parameters from the Sub-Threshold,Low-Noise regime. The
%   target spike train has ISIs that are exponentially distributed with a
%   mean of $1.5$ a.u. There are $N=20$ spikes and $M=10$ realizations of the
%   control strategy}
%   \label{fig:targettrain_cl_lownoise}  
% \end{center}
% \end{figure}   
% \begin{figure}[htp]  
% \begin{center}     
%   \includegraphics[width=.99\textwidth]{Figs/TrainController/SUBTLN_ol_skip_trains_sim_10.pdf}
%   \caption[ ]{SKIPPING MISSED TARGET. How good the {\sl open-loop} controller is at hitting a target
%   spike train. Uses the parameters from the Sub-Threshold,Low-Noise regime. The
%   target spike train has ISIs that are exponentially distributed with a
%   mean of $1.5$ a.u. There are $N=20$ spikes and $M=10$ realizations of the   
%   control strategy}
%   \label{fig:targettrain_ol_lownoise}
% \end{center}
% \end{figure}
% \begin{figure}[htp]  
% \begin{center}     
%   \includegraphics[width=.99\textwidth]{Figs/TrainController/SUBTHN_cl_skip_trains_sim_10.pdf}
%   \caption[ ]{SKIPPING MISSED TARGET. How good the {\sl closed-loop} controller is at hitting a target
%   spike train. Uses the parameters from the Sub-Threshold,High-Noise regime. The
%   target spike train has ISIs that are exponentially distributed with a
%   mean of $1.5$ a.u. There are $N=20$ spikes and $M=10$ realizations of the
%   control strategy}
%   \label{fig:targettrain_cl_highnoise}
% \end{center}
% \end{figure}  
% \begin{figure}[htp]  
% \begin{center}     
%   \includegraphics[width=.99\textwidth]{Figs/TrainController/SUBTHN_ol_skip_trains_sim_10.pdf}
%   \caption[ ]{SKIPPING MISSED TARGET. How good the {\sl open-loop} controller is at hitting a target
%   spike train. Uses the parameters from the Sub-Threshold,High-Noise regime. The
%   target spike train has ISIs that are exponentially distributed with a
%   mean of $1.5$ a.u. There are $N=20$ spikes and $M=10$ realizations of the
%   control strategy}
%   \label{fig:targettrain_ol_highnoise}
% \end{center}
% \end{figure}  
\clearpage
\section{Conclusions / Discussion} 
We have analyzed and computed the optimal control of spikes in two scenarios -
one where the underlying voltage of the neuron is observable to the controller
(closed-loop control) and one where only the spikes are observable (open-loop
control).

ALEX: SUBSECTIONS ARE FOR DRAFT ONLY - THEY WILL BE FUSED
TOGETHER IN FINAL SUBMISSION
\subsection{different regimes}
Naturally, the ability to control the system is most clearly affected by the
level of the noise. In a sense, the noise acts like an adversary - in our
context it has no beneficial role, but to obstruct precise spiking.

The bias current has a more helpful role at least in our examples, where a
high value of the bias tends to help precise spiking. Of course, where it helps
or hinders depends on the value of the desired spike time - a combination of a
high positive bias and a 'distant' spike time, will tend to be difficult to
control as the system will naturally tend to spike earlier than desired.

\subsection{computational matters}
In both cases finding the optimal control is computationally intensive as we
numerically solve partial differential equations. If these algorithms were to be
practical in an online setting, some more work would have to be done in order to
ensure they can be computed very quickly (in the millisecond range) or that
they can be pre-computed.

\subsection{objective criterion}
As our criterion to optimize, we have chosen the squared deviation between the
realized time and some {\sl a priori} chosen time. We have done so as it
appeared to us the simplest objective, representing the idea. Other
objectives such as the absolute value of the difference, $|\ts - \T|$ or its
exponential, $\exp(|\ts - \T|)$, are directly applicable. A different approach
may be to require that $\Exp[\ts] = \T$. It is non-trivial to do this in the
dynamic programing framework, while it is fairly straightforward while working
with the forward equation in the open-loop context. A similarly comment may be
made about maximizing the
likelihood of $\ts = T$, which is the objective chosen in \cite{Ahmadian2011}.

\subsection{relation to other works}
Our work has strayed pretty close to the paradigm of \cite{Ahmadian2011} in
trying to make the neuron spike at a particular time. Other goals have been
formulated, most notably the idea of desynchronizing a population of neurons,
\cite{Nabi2011}, where it is not so important when the neurons spike as long as
they do not spike at the same time. We have also incorporated into our objective
the minimization of total energy used to achieve our goal, which as discussed in
\cite{Ahmadian2011} is sensible given the potential damage on the cell of
accumulating charge. Furthermore, we have assumed that our control is
constrained in magnitude, which is natural given physical limitations on
equipment and safety considerations. Another possible constraint on the control
is that it is charge-balanced, meaning that its time integral is zero, $\int u
\intd{t} = 0$. Such constraints are most easily posed in the context of
deterministic spiking models like the Phase-Response Curve, see e.g. Danzl et
al. \cite{Danzl2010}. It is possible to add them to the open-loop control, since
the control is still deterministic, but it is not-trivial. It is even less
trivial when one uses Dynamic Programing \{ALEX: IT IS WORTH THINKING ABOUT, AS
IT MAKES FOR A VERY COOL MATH PROBLEM, BUT WHAT IS THE PAY-OFF?\}. In
particular, insisting on charge-balance in our schemes will make it much more
difficult to apply the Terminal Conditions used in deriving both the closed- and
open-loop controls.

\subsection{novelty}
We have developed a stochastic optimal control scheme which is not limited
by the intensity of the noise in the system. Given this scope, the
numerical demands on our scheme are non-trivial, but our experience and
experiments show that these demands are potentially feasible. It
however remains to be seen whether they can be successfully applied in the lab
and beyond.

\subsection{{\sl Stochastic} effects} 
We show how wrong the control strategy can be if we assume that the noise is
negligible.

\appendix
\section{Calculating the Terminal Conditions for the HJB equation}
\label{sec:valuefun_TCs}
Here we explain how to calculate the first two moments of the time to-spike,
$\Exp \big[\trem^p \,\Big|\, X_\T = x, \a(t) = \amax \big].$, which is needed in setting the terminal conditions for both,
$\v(x,\T)$ and $p(x,\T)$, see \cref{eq:OC_LS_HJB_full} and \cref{eq:adjoint_pde_OU}.

Let 
\begin{equation} 
\begin{array}{lcl}
\Tone(x_0) &=& \Exp \Big[\trem \,\Big|\, X_\T = x, \a(t) = \amax
\Big]
\\
\Ttwo(x_0) &=&
\Exp \Big[\trem^2 \,\Big|\, X_\T = x, \a(t) = \amax \Big],
\end{array}
\end{equation}
where as before, $\t$ is the remaining time-to-spike in
\cref{eq:X_evolution_uo}, at the desired spike-time $t=\T$, given $X_\T = x_\T$
and $\a(t) = \amax$.
Naturally, the $\Ti$'s depend on $x_\T$ and are affected by the value of the
various parameters, $\a,\b,\tc$.

Actually, we only  need $\Ttwo$ for the TCs in \cref{eq:OC_LS_HJB_full} and \cref{eq:adjoint_pde_OU}. 
but $\Tone$ is needed as an auxiliary.

Indeed, analytical expressions exist for $\Ti$, e.g.\ see \cite{Inoue1995}.
However in our experience, they are awkward to work with numerically, especially
as they involve infinite alternating sums which can easily overflow in numeric
calculations. Moreover, they are not particularly convenient if you need $\Ti
(x_0)$ for a wide range of $x_0$'s as we do here, since they calculate the value
independently for each $x$. Instead we will follow and slightly extend Jacobs\footnote{Jacobs only discusses how to calculate the first moment
$\Tone$, however it is quite clear from his discussion how to proceed
recursively to obtain any higher moment $\Ti$, by using previously calculated
lower moments. This is what we do here. Moreover the relation between mean exit
times and boundary value problems is discussed in most
introductory books on stochastic processes, see \cite{Oksendal2007,Evansa} etc.},
\cite{Jacobs}, in deriving and solving a differential equation for
$\Ti(x_0)$.

Basically we are calculating the mean-time-squared to exit an interval $[\xmin,
\xth]$. Theoretically, $\xmin = -\infty$, but for our purposes we will set it to
some finite value and impose reflecting boundaries there - this is a very good
approximation for $\xmin \ll 0$, as long as $\a = \amax >0$, which is what we
have.

Let $B(y,t|x) = \Prob[X_0 = y| X_t = x]$ for a fixed $x$. Then $B$ solves the
following (backward Kolmogorov) PDE
\begin{equation}
-\di_t B = (\a - \frac y\tc)\di_yB - \frac {\b^2} 2 \di_y^2 B
\label{eq:backward_Xdensity}
\end{equation}

Let $\G(t,y)$ be the survival function for $T$ for $X_0 = y$. 
$$\G(t,y) = \Prob[T>t | X_0 = y] = \Prob[X_{s\leq t} \in [\xmin,
\xth] | X_0 = y]$$.

Because the dynamics do not depend explicitly on time, we have:
$$
\G(t,y) = \int_\xmin^\xth B(y,t|x) \intd{x}
$$
and so $\G$, being a definite integral of $B$ satisfies the same PDE:
\begin{equation}
-\di_t \G = (\a - \frac y\tc) \di_y\G - \frac {\b^2} 2 \di_y^2 \G
\label{eq:backward_SDF}
\end{equation}

With the BCs, ICs:
\begin{equation}
\begin{cases}
\G(t,\xth) = .0 & \quad (T = t \implies \Prob[T>t] = 0)
\\
\di_x \G(t,\xmin) = .0  & \quad (\textrm{due to reflecting boundary for X})
\\
\G(0, y) = 1. & \quad (T \textrm{ is definitely} > 0 \textrm{ for } x_0 \in
(\xmin, \xth) )
\end{cases}
\end{equation}

If we solve the PDE for $\G$ over $t \in [0,\infty)$, then we can calculate
$\Ttwo(x_0)$ as:
\begin{eqnarray}
\Ttwo(x_0) &=& \int_0^\infty t^2 \cdot (-\di_t\G(t,x_0)) \intd{t}
\\
		   &=& \int_0^\infty 2t \cdot \G(t,x_0) \intd{t}
\end{eqnarray}
assuming it exists.

Since $\Ttwo$ is an integral over time, we can reduce the PDE in
\cref{eq:backward_SDF} for the distribution to a BVP for the moment(s) by
integrating over time. Here is how this is done: multiply both sides  of
\cref{eq:backward_SDF} by $2t$, and integrate with respect to time:
\begin{equation}
\int_0^\infty 2t (-\di_t \G) \intd{t}
= 
\int_0^\infty  2t \left[ (\a - \frac y\tc) \di_y\G - \frac {\b^2} 2
\di_y^2 \G\right]
\intd{t}
\end{equation}
The LHS turns out to be $2\Tone$, so 
\begin{equation}
2\Tone 
=(\a - \frac y\tc)   \di_y \Ttwo 
- \frac {\b^2} 2
\di_y^2 \Ttwo
\label{eq:BVP_Ttwo}
\end{equation}
Similarly we can derive an equation for $\Tone$, namely:
\begin{equation}
1
=(\a - \frac y\tc)   \di_y \Tone 
- \frac {\b^2} 2
\di_y^2 \Tone
\label{eq:BVP_Tone} 
\end{equation}
In both cases, the BCs will be:
\begin{equation}
\begin{cases}
\Ti(\xth) = .0 & \quad (\textrm{we are already spking})
\\
\di_x \Ti(\xmin) = .0  & \quad (\textrm{due to reflecting boundary for }X)
\end{cases}
\label{eq:BVP_Ti_BCs}
\end{equation}
Equations \ref{eq:BVP_Tone} and \ref{eq:BVP_Ttwo} can be reduced to first order
differential equations and solved analytically, at least up to some definite integral,
but the analytical solution is not particularly useful or necessary for
numerical calculations. Instead, we opt to calculate the solutions to
\ref{eq:BVP_Tone} and \ref{eq:BVP_Ttwo} directly via an ODE solver, starting
from $T_i(\xth) = .0$ and integrating down for $x  \in [\xmin, \xth]$.
That is how we have obtained the terminal conditions for both $\v(x,\T)$ and
$p(x,\T)$ in our calculations above.

\bibliographystyle{plain}
% \bibliographystyle{unsrt}
% \bibliographystyle{bmc_article}
% \section*{References}  \bibliographystyle{iopartnum}   
\bibliography{library,local}

\end{document}