### 2013-08-22 phd_thesis_plan.tex

\documentclass{article}

\usepackage{amsmath}
\usepackage{amscd}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{amsthm} 
\usepackage{amsfonts}
\usepackage{amsthm}

\usepackage{circuitikz}
\usepackage{pgf}
\usepackage{tikz}
\usetikzlibrary{arrows,snakes,backgrounds}
% \usetikz
\usepackage{subfig}

\usepackage[super]{nth}
% \usepackage{appendix}
% \usepackage{listings}
% \usepackage{color}

\usepackage{hyperref}
%\usepackage{url}

\usepackage{cleveref}

\usepackage{aviolov_style}  
\usepackage{local_style}

\begin{document}

\title{Math Thesis Plan}
\author{alexandre iolov\\ $<$\href{mailto:aiolo040@uottawa.ca}
		{aiolo040 at uottawa dot ca}$>$
		}

\date{\today}

\maketitle

\tableofcontents

\section{Thesis Proposal}
I propose we call the thesis:

{\large \sl ``Estimation and Control in Single Neural Models''}

I propose that the thesis consists of three distinct sections

\begin{enumerate}
  \item Our estimation scheme for the sinusoidal LIF problem 
  \item The spike time optimal control for a noisy LIF 
  \item A discustion on Optimal Design which combines elements from the two
  studies above. That is the control is used to obtain better estimates.
\end{enumerate}

\section{Timelines / Milestones}
I would like to submit my thesis ideally by {\sl January 22nd, 2014}. This is
the last day to do so without signing up for the winter semester.

Failing that, I would aim for {\sl May 22nd, 2014}

What kind of milestones / schedule would make it likely that one of the
above happens?

\end{document}